Exploring how changes in cause of death have narrowed the black-white life

### 2017-02-28 Manuscript/Manuscript_COD_decomp.Rmd
added cross sectional decomp 2013

age_cod_results$COD <- factor(age_cod_results$COD, levels(age_cod_results$COD)[c(3, 2, 4, 6, 5, 1)])
age_decomp_results$age <- factor(age_decomp_results$age, levels = levels(age_decomp_results$age)[c(1, 2, 11, 3:10, 12:19)])
age_cod_results$age <- factor(age_cod_results$age, levels = levels(age_cod_results$age)[c(1, 2, 11, 3:10, 12:19)])

@@

>>>>

age_cod_results$COD <- factor(age_cod_results$COD, levels(age_cod_results$COD)[c(3, 2, 4, 6, 5, 1)])
cod_decomp_results$COD <- factor(cod_decomp_results$COD, levels(cod_decomp_results$COD)[c(3, 2, 4, 6, 5, 1)])
age_decomp_results$age <- factor(age_decomp_results$age, levels = levels(age_decomp_results$age)[c(1, 2, 11, 3:10, 12:19)])
age_cod_results$age <- factor(age_cod_results$age, levels = levels(age_cod_results$age)[c(1, 2, 11, 3:10, 12:19)])

@@

**Fig 7: Decomposition of the male black-white life expectancy gap by cause of death in 2013 by US state**

```{r cross-sectional-decomp-males}
cod.2013.males <- cod_decomp_results %>% filter(year == 2013 & sex == "Male") %>% 
  mutate(widened_gap = sign(COD_cont_yrs_mean), state.n = as.numeric(state))

y.labels.2013 <- levels(factor(cod.2013.males$state))
y.num.breaks.2013 <- length(levels(factor(cod.2013.males$state)))

cod.2013.males <- make_pretty_decomp_plot(decomp.table = cod.2013.males, 
                                          strat.var.1 = "state",
                                          sign.var = "widened_gap",
                                          decomp.var = "COD_cont_yrs_mean",
                                          decomp.var.prop = "COD_cont_prop_mean",
                                          partition.bar.var = "COD",
                                          type.of.data = "practice")

males.2013 <- ggplotly(
  ggplot(cod.2013.males, aes(y = state.n, x = start)) +
    geom_rect(aes(xmin = start, ymin = state.n - 0.45, 
                  ymax = state.n + 0.45, 
                  xmax = finish, fill = COD), 
              color = "white") +
     scale_y_continuous(breaks = 1:y.num.breaks.2013, labels = y.labels.2013) +
    theme_minimal())  %>% layout(xaxis = list(title = "Contribution to gap, years"), yaxis = list(title = NA, autorange = "reversed"))

males.2013
```

**Fig 8: Decomposition of the female black-white life expectancy gap by cause of death in 2013 by US state**

```{r cross-sectional-decomp-females}
cod.2013.females <- cod_decomp_results %>% filter(year == 2013 & sex == "Female") %>% 
  mutate(widened_gap = sign(COD_cont_yrs_mean), state.n = as.numeric(state))

y.labels.2013 <- levels(factor(cod.2013.females$state))
y.num.breaks.2013 <- length(levels(factor(cod.2013.females$state)))

cod.2013.females <- make_pretty_decomp_plot(decomp.table = cod.2013.females, 
                                          strat.var.1 = "state",
                                          sign.var = "widened_gap",
                                          decomp.var = "COD_cont_yrs_mean",
                                          decomp.var.prop = "COD_cont_prop_mean",
                                          partition.bar.var = "COD",
                                          type.of.data = "practice")

females.2013 <- ggplotly(
  ggplot(cod.2013.females, aes(y = state.n, x = start)) +
    geom_rect(aes(xmin = start, ymin = state.n - 0.45, 
                  ymax = state.n + 0.45, 
                  xmax = finish, fill = COD), 
              color = "white") +
     scale_y_continuous(breaks = 1:y.num.breaks.2013, labels = y.labels.2013) +
     #scale_x_continuous(breaks = seq(-2, 16, by = 1), labels = seq(-2, 16, by = 1)) +
    theme_minimal())  %>% layout(xaxis = list(title = "Contribution to gap, years"), yaxis = list(title = NA, autorange = "reversed"))

#ggplot(data = cod.2013.females, aes(x=state, y=COD_cont_yrs_mean, fill=COD)) + geom_bar(stat = "identity", col = "white") + coord_flip() + theme_minimal()

# p <- plot_ly(cod.2013.females, x = ~state, y = ~COD_cont_yrs_mean, type = 'bar', name = 'blah') %>%
#   layout(yaxis = list(title = 'Count'), barmode = 'stack')

# 
# info <- plotly_build(females.2013)
# str(info)
# names(info$x$data[[1]])
```

**Figure 9: Distribution of the contribution of each cause of death to the life expectancy gap in males, 2013**

```{r}
ggplot(cod.2013.males, aes(COD_cont_yrs_mean)) + 
  geom_histogram(aes(fill = COD), col = "white") + 
  theme_minimal() + 
  geom_vline(xintercept = 0) +
  facet_wrap(~COD) + 
  ylab("Number of states") + guides(fill = F) +
  scale_y_continuous(breaks = c(seq(from = 0, to = 15, by = 5)), minor_breaks = c(seq(from = 0, to = 20, by = 5))) +
  xlab("Contribution to gap in 2013 (years)") + scale_x_continuous(breaks = c(-1:4), minor_breaks = c(-1:4))
```

**Figure 10: Distribution of the contribution of each cause of death to the life expectancy gap in females, 2013**

```{r}
# cod.2013.females$CI.width <- cod.2013.females$COD_cont_yrs_ucl - cod.2013.females$COD_cont_yrs_lcl
# ggplot(cod.2013.females, aes(y = COD_cont_yrs_mean, x = CI.width)) + 
#   geom_point(aes(col = COD), alpha = 0.8) + geom_text(aes(label = stabbrs), hjust = 1.5, vjust = 0.1, size = 3) +
#   facet_wrap(~COD, scales = "free_x") + 
#   geom_abline(intercept = 0 , slope = 0) +
#   ylab("Contribution to gap in 2013 (years)") + xlab("Width of credible interval")

ggplot(cod.2013.females, aes(COD_cont_yrs_mean)) + 
  geom_histogram(aes(fill = COD), col = "white") + 
  theme_minimal() + 
  geom_vline(xintercept = 0) +
  facet_wrap(~COD) + 
  ylab("Number of states") + guides(fill = F) +
  scale_y_continuous(breaks = c(seq(from = 0, to = 15, by = 5)), minor_breaks = c(seq(from = 0, to = 20, by = 5))) + 
  xlab("Contribution to gap in 2013 (years)") + scale_x_continuous(breaks = c(-1:4), minor_breaks = c(-1:4))
```
